Compositional Quantitative Resource Analysis of C Programs

### 2014-06-27 paper/main.tex
start soundness of the analysis

>>>>

\paragraph{Soundness Proof.}

The soundness of the analysis leverages the quantitative logic.
We propose to translate judgements of~\pref{fig:auto} into
quantitative triples.  We define a translation function
$\tr$, such that if a judgement $J$ in the automatic analysis
is derivable, $\tr(J)$ is derivable in the quantitative
logic.  Using this translated judgement and the soundness
of our logic it is possible to get a certified resource bound
for the analyzed program.

The translation of an assertion $(\Gamma, Q)$ in the automatic
analysis is defined by
$$
  \tr(\Gamma, Q) := \lambda\state .\, (\Gamma(\state)) + \Phi_Q(\state),
$$
where we write $\Phi_Q$ for the unique linear potential function
defined by the quantitative annotation $Q$.  We also need to
translate the assumptions in function contexts of the automatic
analysis, this is done as follows
%
\begin{align*}
\tr(\Gamma_f, Q_f, \Gamma'_f, Q'_f) :=
&\forall z \, \vec v \, v . (
  \lambda \_ \, \vec v \, \state .\, \tr(\Gamma_f, Q_f)(\state[\Vargs \mapsto \vec v]), \\
&\phantom{\forall z \, \vec v \, v . (}
  \lambda \_ \, v \, \state .\, \tr(\Gamma'_f, Q'_f)(\state[\Vret \mapsto v])
  ).
\end{align*}
%
These definitions let us translate the judgement $J = B; R; P \vdash
S \dashv P'$ in the context $\Delta$ by
$$
  \tr(J) :=
  (\lambda f .\, \tr(\Delta(f))); \tr(B); \tr(R) \vdash
    \htriple{\tr(P)}{S}{\tr(P')}.
$$
The soundness of the automatic analysis can now be stated
formally with the following theorem.
%
\begin{theorem}[Soundness of the analysis]
  If $J$ is a judgement derived with the rules of
  \pref{fig:auto}, then $\tr(J)$ is a quantitative
  Hoare triple derivable with the rules of
  \pref{fig:logic}.
\end{theorem}
%
\noindent
The proof of this theorem is constructive and basically
maps each rule of the automatic analysis directly to
its counterpart in the quantitative logic.  With the
computational content of the proof, we have an
effective algorithm to construct proof objects for
upper bounds derived automatically.
